% !TeX root = ../Masterarbeit.tex

\documentclass[Masterarbeit.tex]{subfile}

\begin{document}
	
	\chapter*{Abstract}
	
	Slowly varying ocean heat content is one of the most important variables when describing climate variability on interannual to decadal time scales. Since observation-based estimates of ocean heat content require extensive observational coverage, incomplete observations are often combined with numerical models via data assimilation to simulate the evolution of oceanic heat. However, incomplete observations, particularly in the subsurface ocean, lead to large uncertainties in the resulting model-based estimate. As an alternative approach, \cite{kadow_artificial_2020} have proven that artificial intelligence can successfully be utilized to reconstruct missing climate information for surface temperatures. In the following, I investigate the possibility to train their three-dimensional convolutional neural network to reconstruct missing subsurface temperatures to obtain ocean heat content estimates with a focus on the North Atlantic ocean. \\
	The network is trained and tested to reconstruct a 16 member Ensemble Kalman Filter assimilation ensemble constructed with the Max-Planck Institute Earth System Model for the period from 1958 to 2020. Specifically, I examine whether the partial convolutional U-net represents a valid alternative to the Ensemble Kalman Filter assimilation to estimate North Atlantic subpolar gyre ocean heat content. Additionally, I draw conclusions from this reconstruction in regard to the assimilation's uncertainty outside of its ensemble spread. \\
	The neural network is capable of reproducing the assimilation reduced to datapoints with observational coverages within its ensemble spread with a correlation coefficient of 0.93 over the entire time period and of 0.99 over 2004 -- 2020 (the Argo-Era). Additionally, the network is able to reconstruct the observed ocean heat content directly from observations for 12 additional months with a correlation of 0.97, essentially replacing the assimilation experiment by an extrapolation. When reconstructing the pre-Argo-Era (1958 -- 2004), the network's training data still consists only of data assimilations from the Argo-Era, which are assumed to possess lower uncertainties due to more observational input. The lower correlation in the resulting reconstruction of masked assimilations from the pre-Argo-Era indicates significant uncertainties in the assimilation outside of its ensemble spread at times with low observational density. My results therefore demonstrate that the neural network is not only capable of reproducing the observed ocean heat content over its training period, but also before and after as well as improving estimates of the assimilation's uncertainty making it a suitable candidate to step-wise extend or replace data assimilation. \\
	
	
\end{document}