% !TeX root = ../Masterarbeit.tex

\documentclass[Masterarbeit.tex]{subfile}

\begin{document}
	
	\chapter{Summary and Conclusions}
	
	This study aims at examining if neural networks, more specifically partial convolutional U-nets, are able to reproduce or replace EnKF data assimilations of SPG OHC. Before drawing any conclusions from the analysis described in section \ref{results}, its main results are revisited and summarized. \\
	While training with Argo era masks on Argo era EnKF assimilations, the partial convolutional U-net described in section \ref{U-net} is able to reconstruct masked assimilations from the Argo era well within the assimilation's ensemble spread. The timeseries of SPG OHC produced by the neural network reproduces even the monthly variability of the reconstructed assimilation member. Furthermore, a look into the OHC maps shows that the neural network is able to reconstruct shows regional OHC patterns and even small scale structures. This assessment is supported by a mean pattern correlation of ?? over the entire Argo era. \\
	When these results are extended to the pre-Argo era, the agreement between assimilation and neural network reconstruction sinks significantly. The neural network reconstruction and assimilation disagree strongly during during the first 10 years. Afterwards, the neural network still shows a warm bias compared to the assimilation. The pattern correlation between both OHC estimates is also significantly lower during the pre-Argo era compared to the Argo era, especially during before 1970. Nonetheless, the neural network reconstruction stays within the assimilation ensemble spread for most of the pre-Argo era. The remaining differences can be interpreted as indicators of uncertainty within the assimilation due to missing observations. Most notably, the AOC's north west corner is misrepresented in the assimilation unlike in the neural network reconstructions when no observations are available, which could explain the warm bias of the neural network during the first half of the pre-Argo era. \\
	The neural network's direct reconstruction of observations shows even larger differences to the assimilation. Here, even during the Argo-era significant differences arise. These differences can be explained by the bias inherent to the assimilation at the points of observation. In addition to a bias, the pattern correlation between assimilation and direct reconstruction is significantly lower than that of the indirect reconstruction. The differences in pattern correlation as well as overall OHC are especially strong during the SPG OHC peak around 2009. Both occurrences are accompanied by a stronger bias and lower pattern correlation of the assimilation compared to the observations. Thus, the neural network seems to transfer differences in its masked input to its overall OHC estimate. That impression is supported by the EN4 objective analysis, which shows a cool bias towards the assimilation during the Argo era similar to that of the neural network reconstruction. \\
	Finally, an analysis of the computational resources used by both the EnKF assimilations and the neural network is conducted. This examination shows that the difference in the assimilation of an additional timestep is insignificant compared to its application. The generation of an entire 16 member ensemble then already shows much higher differences, with the assimilation requiring approximately double the OCT and four times the node hours compared to the neural network. However, the most significant differences occur, when the neural network's ability to reconstruct independent of another training instance is utilized. Thus, the generation of another ensemble member or the reconstruction of adjusted observational data show differences in node hours of up to a factor 1000. \\
	
	Now, these results can be evaluated in concern to this study's research questions.\\
	Given the close agreement between the neural network reconstruction and the EnKF assimilation it can be concluded that partial convolutional U-nets are able to reproduce EnKF assimilations in the Argo as well as pre-Argo era. Especially the comparison to other OHC estimates, such as the shown IAP and EN4 analyses, show that the neural network can reproduce its training input much closer than its uncertainty in the comparison of different mapping methods.\\
	Additionally, the results show that the neural network reconstructions are also able to replace the EnKF assimilation in their function as a OHC reanalysis for new timesteps. Even monthly variations in timesteps outside of its training data are reproduced by the neural network when directly reconstructing observations from the Argo era onwards. \\
	Besides replacing the assimilations, this study has also shown that neural network reconstructions can take a role in estimating their uncertainty outside of the assimilation's ensemble spread. The neural network was able to transfer assimilation patterns from its well observed training period to earlier time spans. Thus, it can be concluded that partial convolutional U-nets can also be utilized to detect uncertainties and inconsistencies in OHC assimilations. Furthermore, the results of the direct reconstruction of observations are highlighting inconsistencies between assimilation and observations at the observed grid points. Therefore, partial convolutional U-nets can also help estimate the uncertainty stemming from biases and differences between assimilation and observations at the observed gridpoints.\\
	Finally, the comparison of computational requirements between both methods show that neural network have the potential of significantly speeding up assimilation of individual variables. Especially the flexible buildup of neural networks can speed up specific functions related to data assimilation, such as the generation of new ensemble members. With growing resolution and larger datasets, this difference between the computational requirements is bound to grow even further \cite{brune_personal_2022}. \\
	
	In general, this study has shown that neural networks have the potential to significantly impact data assimilation of ocean variables in two ways. On the one hand, neural networks have the ability to reproduce mapping methods. This ability is highlighted by the potential for improvement discussed in section \ref{discussion} and the accuracy to which this partial convolutional U-net is already able to reproduce its training data. On the other hand, they are capable of giving insights into the uncertainty of other assimilation techniques, especially for undersampled regions and variables. Ultimately, this study has once again shown that it only represents one of the first steps towards machine learning greatly impacting data assimilation and the climate sciences in general.\\
	
	
	
	summary:
	argo
	preargo
	future
	direct
	uncertainties
	resources
	
	conclusions:
	reproducing?
		argo
		perargo
		future
	replacing?
		argo
		preargo
		future
	more efficient?
		uncertainties in assimilation?
	from masked
	from direct

	
	
	
	
	
\end{document}